How to improve your relationship with your future self.

### 2016-06-24 workflow.Rmd
Update workflow.Rmd
fixed an error, will continue soon!
made some changes, not done yet, but loving it!

How long does it take from planning a study to publication and to the first reproduction of the study after publication? Is three years too short? Is ten years too long? We suspect that few of our colleagues in the social and behavioral sciences conceive of, field, write, and publish a study in faster than about three years. We also suspect that, if some other scholar decides to re-examine the analyses of a published study, that it will occur after publication, and that this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis. Othertimes, this process is organized more broadly (cite IIIE program, cite Open Science Foundation on psychology).]

If we cannot count on our memories about why we made such and such a design or analysis decision, then what should we do? How can we minimize regret with our past decisions? How can we improve our relationship with our future self?  This essay is a heavily revised version of @bowers2011tpm and provides some suggestions for practices that will make such reproduction occur much more easily and quickly in the event that famous papers require special scrutiny.  Specifically, this piece aims to amplify some of what we already ought to know,  and to add to some of those ideas
given current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists. Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with the DA-RT inititiave <http://www.dartstatement.org/> @lupia2014openness. These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations.] This essay is an attempt at providing practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on other bad past decisions but not on decisions about data analysis and the production of scholarly papers.

@@

All results (numbers, comparisons, tables, figures) should arise from
code, not from a series of mouse clicks or copying and pasting. If I wanted
to re-create the figure from 2011 but include a new variable or
specification, I should be able to do so with just a few edits to the
code rather than knowledge of how I (or you) used a pointing device in
your graphical user interface some years ago.

Using R [@R:2011], for example, I might specify that the file `fig1.pdf` was
produced by the following commands in a file called `fig1.R`.^[The command
`please-plot`and some other R functions used in this essay come from the
`MayIPleaseDoStatistics` package which emphasizes politeness in data analysis.
Functions like `please-plot`can be blocked and more polite versions such as
`may-I-please-have-a-plot`can be required using `options(politeness=99)`. We do
not recommend setting negative values for the politeness option should you use this package.]

@@

Now, in the future if I wonder how 'that plot on page 10' was
created, I will know: (1) 'that plot' is from a file called
`fig1.pdf`and (2) `fig1.pdf`was created in
`fig1.R`. In a future where R still exists, changing the figure
will require quick edits of commands already written.  In a future
where R does not exist, I will at least be able to read the plain text
R commands and use them to write code in my new favorite statistical
computing language: R scripts are written in
[plain text](http://en.wikipedia.org/wiki/Plain_text), and plain
text is a format that will be around as long as computer programmers
write computer programs.^[Since R is open
  source, I will also be able to download an old version of R, download an
  old-fashioned open-source operating system (like Ubuntu 10), and
  run the old-fashioned statistical computing environment in the
  old-fashioned operating system in a virtual machine on my
  new-fashioned actual machine.]

Moreover, realize that *file names send messages* to your future
self. This means that if you name your files with evocative and descriptive names, your
collaborators are less likely to call you at midnight asking for help and you will remove some regret from your future self. For example, if you are studying inequality and protest, you might try naming a file something like  `inequality-and-protest-figures.R` instead of  `temp9`or
`supercalifragilisticexpialidocious`.  By the way, the extension
`.R` tells us and the operating system that the file contains R
commands. This part of the filename enables us to quickly search our
antique hard drives for files containing R scripts.

**Step 1** If we know the provenance of results,
future  or current collaborators can quickly and easily reproduce
and thus change and improve upon the work.

@@

Data analysis involves a long series of decisions. Each decision
requires justification.  Some decisions will be too small and
technical for inclusion in the published article itself. These need to
be documented in the code itself [@nagler1995coding]. Paragraphs
and citations in the publication will justify the most important
decisions. So, one must code to communicate with yourself and
others. There are two main ways to avoid forgetting the reasons you
did something with data: comment your code and tightly link your code
with your writing.^[One can also try the R command
  `put-it-in-my-mind(reason,importance="high")` to firmly place a
  reason for a decision into the mind of the analyst. I myself have
  not had much luck with this function.]

>>>>

How long does it take from planning a study to publication? And subsequently, how much time passes to the first reproduction of a study after publication? Is three years too short? Is ten years too long? We suspect that few of our colleagues in the social and behavioral sciences conceive of, field, write, and publish a study in faster than about three years. We also suspect that, if some other scholar decides to re-examine the analyses of a published study, it will occur after publication, and that this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis. Other times, this process is organized more broadly (cite IIIE program, cite Open Science Foundation on psychology, cite the Camerer 2016 Science paper?, cite the cool "Crowdsourcing Analysis" project by Silberzahn and Uhlmann Nature 2015?).]

If we cannot count on our memories about why we made such and such a design or analysis decision, then what should we do? How can we minimize regret with our past decisions? How can we improve our relationship with our future self?  This essay is a heavily revised and updated version of @bowers2011tpm and provides some suggestions for practices that will make such reproduction occur much more easily and quickly in the event that famous papers require special scrutiny.  Specifically, this piece aims to amplify some of what we already ought to know, and to add to some of those ideas given current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists. Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with the DA-RT inititiave <http://www.dartstatement.org/> @lupia2014openness. These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations.] This essay is an attempt at providing practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on other bad past decisions but not on decisions about data analysis and the production of scholarly papers.

@@

All results (numbers, comparisons, tables, figures) should arise from
code, not from a series of mouse clicks or copying and pasting. If you wanted
to re-create the figure from say 2011 but include a new variable or
specification, you should be able to do so with just a few edits to the
code rather than knowledge of how you used a pointing device in
your graphical user interface all those years ago.

For example, using R [@R:2011], a open-source statistical software program, I might specify that a file, called `fig1.pdf` was
produced by the following set of commands in a file called `makefig1.R`.^[The command
`please-plot`and some other R functions used in this essay come from the
`MayIPleaseDoStatistics` package which emphasizes politeness in data analysis.
Functions like `please-plot`can be blocked and more polite versions such as
`may-I-please-have-a-plot`can be required using `options(politeness=99)`. We do
not recommend setting negative values for the politeness option should you use this package.]
Lets look at some annotated R code you could produce:

@@

Now, in the future if you wonder how 'that plot on page 10' was
created, you will know: (1) 'that plot' is from a file called
`fig1.pdf`and (2) `fig1.pdf`was created in
`makefig1.R`. Any time in a future (provided R still exists), changing the figure
will require some quick edits of commands already written.  In a future
where R does not exist, you (or someone else) will at least be able to read the plain text
R commands and use them to write code in a new favorite statistical
computing language: R scripts are written in
[plain text](http://en.wikipedia.org/wiki/Plain_text), and plain
text is a format that will be around as long as computer programmers
write computer programs.^[Since R is open
  source, you will also be able to download an old version of R, download an
  old-fashioned open-source operating system (like Ubuntu 10), and
  run the old-fashioned statistical computing environment in the
  old-fashioned operating system in a virtual machine on my
  new-fashioned actual machine.]

Moreover, coding saves time. Often, a lot of time, especially if projects increase (in the number of files, partners, analysis). Manually importing
a data file into R, may be effortless. Importing 100 files is another issue and the costs of each manual action add up quickly.

Also, realize that *file names send messages* to your future self. This means that if you name your files with evocative and descriptive names, your
collaborators are less likely to call you at midnight asking for help and you will remove some regret from your future self. For example, if you are studying inequality and protest, you might try naming a file something like  `inequality-and-protest-figures.R` instead of  `temp9`or
`supercalifragilisticexpialidocious`.  More on this below. By the way, the extension
`.R` tells us and the operating system that the file contains R commands. This part of the filename enables us to quickly search our
antique hard drives for files containing R scripts.

Finally, coding helps us to avoid making mistakes. For example, say in our example above we are interested in how many people protest. We may use a data file containing all protests for several years. We can then create a table with the number of protesters by copying the results manually to a document you use to write your paper. In copying we can make mistakes. So a better approach is to ask R to automatically create the table (in the format you like, with horizontal lines, 3 decimals, a title, etc) and output this in whatever file type you need (.pdf, .rtf, etc). Now when we obtain new data (for additional years, or updated figures for existing data) a new table can be created quickly, so we make less mistakes and save time!

NOTE we could include some code here to provide an example of automation of results?

**Step 1** Code everything that can be coded. If we know the provenance of results, future or current collaborators make less mistakes and can quickly and easily reproduce (and thus change and improve) upon the work.

@@

Data analysis involves a long series of decisions. Each decision
requires a justification.  Some decisions will be too small and
technical for inclusion in the published article itself. These need to
be documented in the code itself [@nagler1995coding]. Paragraphs
and citations in the publication will justify the most important
decisions. So, one must code to communicate with yourself and
others. There are two main ways to avoid forgetting the reasons you
did something with data: comment your code and tightly link your code
with your writing.^[One can also try the R command
  `put-it-in-my-mind(reason,importance="high")` to firmly place a
  reason for a decision into the mind of the analyst. I myself have
  not had much luck with this function.]
